---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
unzip("activity.zip")  #Unzip data.
activity<-read.csv("activity.csv")  ## read file into R
```


## What is mean total number of steps taken per day?
The best way to determine this is to convert days into a factor and make a new dataframe that is the result of summing the step in each day. Then use this dataframe to make a mean over all days.
```{r}
activity$date<-as.Date(activity$date)
activity$date<-as.factor(activity$date)
results<-tapply(activity$steps,activity$date, sum, na.rm=TRUE)
### Datatable "results" keeps the information.
mean(results)
median(results)
hist(results, breaks = 10, main = "Frequecy of days with steps in range")
```
Results show a normal-looking distribution with possible outliers at zero and >20,000. 

## What is the average daily activity pattern?
The only information about time in the day is contained in the interval number. 

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
